Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-01-19 spie/paper.tex
geom stuff amlost

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
our geometry and explains a simple approach towards texturing. Section
\ref{sec:existingApproaches} covers more sophisticated existing
approaches, and their performance on our datasets. Section
\ref{sec:2dAlignment} contains our efficient approach towards image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes our two methods of compositing images. Section
\ref{sec:resultsAndConclusions} contains results and conclusions.

@@

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a simple texturing approach can be performed by discretizing the
target plane into small square tiles, in our case 5 pixels across, and
choosing an image to texture each tile directly.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

@@

Past work with our data acquisition backpack integrated image
stitching with an iterative localization algorithm, and performed
homography-based camera pose refinement. When run on long chains of
images, especially where features are sparse, it produces distorted
textures, as seen in Figure X, due to the problems described
above. This approach is also WHATEVER NONLINEAR, and its iterative
camera adjustment process over our large datasets is not guaranteed to
SOMETHING, and leads to prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift (cite
X,X). Though AutoStitch also performs well in areas with dense
features, it can not handle areas without features, and has trouble
aligning wall sections with even short segments of bare texture. The
example in Figure X was generated after manual tuning, and areas with
fewer visual features or repeating texture patterns simply failed
outright with AutoStitch.

@@

In this section, we describe our method of efficient and robust image
alignment. Though state-of-the-art techniques towards image stitching
work in 3D, we sacrifice THAT WORD for efficiency, and elect to
perform image alignments entirely in 2D. Besides the reduced
complexity, we have found 2D alignment to work well for two main
reasons. First, because we have such a large number of images to
choose from, we have found that the compositing approache in Section X
is able to select a subset of images that align well, given only 2D
alignments. Furthermore, the nature of our input data is such that
localization error chiefly occurs in two dimensions, in the same plane
as the surface being projected onto. Recall that our data acquisition
system is backpack-mounted, with cameras facing to the sides of the
operator. The operator makes efforts to walk parallel to walls, and
the localization and model-generation algorithms that provide our
input conform the operator's path to be straight and parallel to
detected surfaces \cite{kua2012loopclosure, sanchez2012point}. Since
the operator walks in as upright a position as possible, errors in
roll, yaw, and the operator's distance from parallel surfaces are
minimal. Thus, our highest errors stem from uncertainty in the
operator's pitch, equivalent to rotation around the camera axes, as
well as location along surfaces. These equate to 2D rotation and
translation in a surface's plane.

Our alignment procedure consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then matched up with
geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a linear least squares problem in 2D to maximize
matches.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms. Experience
and intuition show that walls in indoor environments often contain
linear features that are either horizontal or vertical, corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. This can be further extrapolated by orienting lines with a
wall's boundaries, for example in areas with a slanted roof.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geomAlign}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) The darker lines are geometry-based lines, while
    lighter lines are lines detected in the image via Hough
    transform. Above is the original projection using the input camera
    poses, and below is the projection after rotation and translation
    for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, the image is rotated and shifted in 2D such that the visual
boundary between two surfaces in an image projection matches the
physical boundary in our digital model. An example of such an
adjustment is in Figure \ref{fig:geometryAlignment}

To perform the rotation and alignment, we use the RANSAC
\cite{fischler1981random} framework. RANSAC allows us to determine an
optimal rotation and translation to match up two groups of line
segments, while accounting for outliers. First, a rotation angle is
computed by gathering all pairs of image line segments and geometry
line segments with less than 20 degrees between them. We then run the
RANSAC algorithm by sampling 4 of these angles at a time, and using
their average as the fitting function. Angles over 2 degrees from this
average are considered outliers. Once this 2d rotation angle is
determined, it is applied to the image, and projections are
re-calculated and image-based lines are re-detected. Now, all pairs of
image line segments and geometry line segments with less than 0.2
degree difference and less than 250 mm distance at their furthest
points are gathered. RANSAC is then similarly employed to calculate
optimal horizontal and vertical translations independently, by
sampling 4 pairs at a time and considering an outlier threshold of 10
mm. If a translation is applied, we indicate that the image was
aligned to geometry in one or both dimensions, which will reduce
further shifting in Section \ref{sec:robustSIFTFeatureMatching}.

>>>>

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
our geometry and explains a simple approach towards selecting images
for texturing. Section \ref{sec:existingApproaches} covers
sophisticated existing approaches to image stitching, and their
performance on our datasets. Section \ref{sec:2dAlignment} contains
our efficient approach towards 2D image alignment, followed by Section
\ref{sec:imageCompositing}, which describes our two methods of
compositing images. Section \ref{sec:resultsAndConclusions} contains
results and conclusions.

@@

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a simple texturing approach can be performed by discretizing the
target plane into small square tiles, and choosing an image to texture
each tile directly.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

@@

Past work with our data acquisition backpack integrated image
stitching with an iterative localization algorithm, and performed
homography-based camera pose refinement. When run on long chains of
images, especially where features are sparse, it produces distorted
textures, as seen in Figure \ref{fig:mosaic3D}(a), due to the problems
described above. This approach is also not closed-form, and its
iterative camera adjustment process over our large datasets leads to
prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift
\cite{panorama2d, autostitch}. Though AutoStitch also performs well in
areas with dense features, it can not handle areas without features,
and has trouble aligning wall sections with even short segments of
bare texture. The example in Figure \ref{fig:mosaic3D}(b) was
generated after manual tuning, and areas with fewer visual features or
repeating texture patterns simply failed outright with AutoStitch.

@@

In this section, we describe our method of efficient and robust image
alignment. Though state-of-the-art techniques towards image stitching
work in 3D, we elect to perform image alignments entirely in
2D. Besides the benifits of reduced complexity, we have found 2D
alignment to work well for two main reasons. First, because we have
such a large number of images to choose from, we have found that the
compositing approaches in Section \ref{sec:imageCompositing} is able
to select a subset of images that align well, given only 2D
alignments. Furthermore, the nature of our input data is such that
localization error chiefly occurs in two dimensions, in the same plane
as the surface being projected onto. Recall that our data acquisition
system is backpack-mounted, with cameras facing to the sides of the
operator. The operator makes efforts to walk parallel to walls, and
the localization and model-generation algorithms that provide our
input conform the operator's path to be straight and parallel to
detected surfaces \cite{kua2012loopclosure, sanchez2012point}. Since
the operator walks in as upright a position as possible, errors in
roll, yaw, and the operator's distance from parallel surfaces are
minimal. Thus, our highest errors stem from uncertainty in the
operator's pitch, equivalent to rotation around the camera axes, as
well as location along surfaces. These equate to 2D rotation and
translation in a surface's plane.

Our 2D alignment procedure consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then rotated and shifted to
match geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a linear least squares problem in 2D to maximize
matches.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms. Experience
and intuition show that walls in indoor environments often contain
linear features that are either horizontal or vertical, corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. This can be further generalized by instead orienting lines
to be parallel with a surface's boundaries, which makes it applicable
to non-rectangular surfaces, such as ceilings, floors, or slanted
walls.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geomAlign}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) In this example, we project two images at once for
    the sake of demonstration, where the red lines are geometry-based
    lines where surfaces intersect, and the green ines are lines
    detected in the image via Hough transform. Above are the original
    projections using input camera poses, and below are the
    projections after rotation and translation for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

Furthermore, at this point in time, image occlusions have not been
accounted for. As a result, some image projections contain texture
that should project to an adjacent surface, sometimes with a visible
linear boundary where the two surfaces meet. If this linear boundary
is detected by Hough transform, the image is rotated and shifted in 2D
such that the visual boundary between two surfaces in an image
projection matches the physical boundary in our digital model. An
example of such an adjustment is in Figure \ref{fig:geometryAlignment}

To perform the rotation and alignment, we use the RANSAC
\cite{fischler1981random} framework. RANSAC allows us to determine an
optimal rotation and translation to match up two groups of line
segments, while accounting for outliers. The parameters we specify
were chosen after some tuning on our datasets, and are likely to vary
based on input data. First, a rotation angle is computed by gathering
all pairs of image line segments and geometry line segments with less
than 20 degrees between them. We then run the RANSAC algorithm by
sampling 4 of these angles at a time, and using their average as the
fitting function. Angles over 1 degree from this average are
considered outliers. Once a 2D rotation angle is determined, it is
applied to the image, and image line segments with length below a
certain threshold are discarded. Now, all pairs of image line segments
and geometry line segments with less than 0.1 degree difference in
orientation and less than 250 mm distance at their furthest points are
gathered. RANSAC is then similarly employed to calculate a shift that maximizes matches among these pairs of lines.

In an ideal case, homography might be applied here, but poop
In our datasets we have success sides for walls, floors and ceilings we have found often difficulty matching, so we often align only in X, and save this info later.
For floors and ceilings, we have more success, and since images smaller upwards, only one match usually, and so we save the slope for later.

optimal
horizontal and vertical translations independently, by sampling 4
pairs at a time and considering an outlier threshold of 10 mm. If a
translation is applied, we indicate that the image was aligned to
geometry in one or both dimensions, which will reduce further shifting
in Section \ref{sec:robustSIFTFeatureMatching}.
